The ChEMBL database as Linked Open Data

### 2013-03-06 article.tex
Note about the nature of this 'ad-hoc' ontology
Fixed the number of assays in the RDF to match the UU end point

However, the previous ChEMBL-RDF conversion used custom predicates and classes to represent RDF triples, defining
an ad-hoc ontology~\cite{Willighagen2011}. The current version, instead, uses community-proposed
ontologies, making the RDF more interoperable with the RDF offerings of different databases. 
The new triple dataset uses standard ontologies
including the Bibliography Ontology~\cite{Giasson2011} and the Citation Typing Ontology (CiTO)~\cite{Shotton2010} for literature
references and citations, and domain ontologies such the Protein Ontology\cite{Natale2011} and the Chemical Information
Ontology\cite{Hastings2011}. Throughout this paper various prefixes are used to denote different ontology 
namespaces to simplify the RDF examples. These prefixes are outlined in Table~\ref{namespaces}.

@@

A similar SPARQL query, replacing \verb+chembl:Target+ with \verb+bfo:MaterialEntity+, returns 1,143,682 entities for which activities are captured in ChEMBL, and by using \verb+chembl:Assay+, the query returns the total of 617,681 ChEMBL assays.
It should be noted that last query returns five entries less than the SQL database, because the
triple generation requires the assay type to be defined, which is not the case for those 
five assays. With \verb+chembl:Activity+ and \verb+bibo:Article+, we count 6,933,068 ChEMBL activities and 44,681 documents,
respectively. The document count is one fewer than appears 
in the ChEMBL 13 release notes~\cite{ChEMBL13ReleaseNotes}, as the result does not include the doc\_id -1, as this refers to the 
unpublished dataset, not itself a valid \verb+bibo:Article+.

>>>>

However, the previous ChEMBL-RDF conversion used custom predicates and classes to represent RDF triples, defining
an ad-hoc ontology~\cite{Willighagen2011}. This implicit ontology reflects the concepts expressed
in the relational database, and are not formally defined. The current version, instead, uses community-proposed
ontologies, making the RDF more interoperable with the RDF offerings of different databases. 
The new triple dataset uses standard ontologies
including the Bibliography Ontology~\cite{Giasson2011} and the Citation Typing Ontology (CiTO)~\cite{Shotton2010} for literature
references and citations, and domain ontologies such the Protein Ontology\cite{Natale2011} and the Chemical Information
Ontology\cite{Hastings2011}. Throughout this paper various prefixes are used to denote different ontology 
namespaces to simplify the RDF examples. These prefixes are outlined in Table~\ref{namespaces}.

@@

A similar SPARQL query, replacing \verb+chembl:Target+ with \verb+bfo:MaterialEntity+, returns 1,143,682 entities for which activities are captured in ChEMBL, and by using \verb+chembl:Assay+, the query returns the total of 617,676 ChEMBL assays.
It should be noted that last query returns five entries less than the SQL database, because the
triple generation requires the assay type to be defined, which is not the case for those 
five assays. With \verb+chembl:Activity+ and \verb+bibo:Article+, we count 6,933,068 ChEMBL activities and 44,681 documents,
respectively. The document count is one fewer than appears 
in the ChEMBL 13 release notes~\cite{ChEMBL13ReleaseNotes}, as the result does not include the doc\_id -1, as this refers to the 
unpublished dataset, not itself a valid \verb+bibo:Article+.
